Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-01-15 5.Bayes.optimality/5.Bayes_optimality_and_minimax.tex
a mess
I tried separating Ch 4 into 2 chapters and I am almost done.  It is still a mess but it might work out nicely!

We establish in this section minimax versions of our results from Section \ref{sec:boundary}.
Specifically, if we restrict ourselves to \emph{the class of thresholding procedures} $\cal T$ (defined in \eqref{eq:thresholding-procedure}), then Bonferroni's procedure is minimax optimal, for \emph{any} fixed dependence structures in the URS class.
This is formalized in Corollary \ref{cor:point-wise-minimax} in Section \ref{subsec:point-wise-minimax}.
We refer to this result as \emph{point-wise} minimax, to emphasize the fact that this optimality holds for every \emph{fixed} URS array.

Meanwhile, if we search over \emph{all procedures}, but expand the space of models to include {\em all} dependence structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization of the finite-sample Bayes optimality of thresholding procedures in Sections \ref{subsec:Bayes-optimality} and \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

@@

\subsection{Point-wise minimax optimality}
\label{subsec:point-wise-minimax}

\begin{proof}[Proof of Corollary \ref{cor:point-wise-minimax}]
The first conclusion \eqref{eq:point-wise-minimax-above} is a restatement of Theorem \ref{thm:sufficient}.

\subsection{Bayes optimality in support recovery problems}
\label{subsec:Bayes-optimality}

\stilian{\fbox{SUGGESTION}}{ What do you think about making the Bayesian stuff a Chapter?  I think it stands well on its own... plus it is somewhat different from the ``non-Bayesian'' setup of the previous support recovery results...}

\stilian{
In studying support recovery problems \citep[e.g.,][]{arias2017distribution}, often the attention is restricted to the class of
thresholding procedures.  This is justified by arguing that such procedures are the ``reasonable'' choice for estimating the support set.
%  In Theorem \ref{thm:sufficient} we saw that a (FWER-controlling) thresholding procedure is indeed consistent for support recovery. 
% It is natural to ask if it will be sufficient to restrict our attention to only such procedures.
We show in this section that, in many cases, the finite-sample Bayes optimal procedures do indeed belong to the class of thresholding procedures.  
Perhaps surprisingly, however, we also show that for general error models, the thresholding procedures are sub-optimal, 
even when the observations are independent.  Roughly speaking, in the additive error 
models \eqref{eq:model-additive},  this distinction depends on whether or not the error density is log-concave 
(Proposition \ref{prop:log-concave} and Corollary \ref{cor:log-convex}).}{\fbox{edited, agree?}}

We shall identify the optimal procedure for support recovery problems under a Bayesian setting with general distributional assumptions 
(including but not limited to additive models \eqref{eq:model-additive}). 
Specifically, consider an ordered set $P = (i_1,\ldots,i_s)$ of $s$ distinct indices in $\{1,\ldots,p\}$, which will be referred to 
as the {\em ordered support} of the signal.  Consider also $s$ (not necessarily equal) signal densities 
$f_{1}, \ldots, f_{s}$ as well as an error density $f_0$. We shall assume that the observations with indices $i_j$ 
are drawn independently from $f_j$:
\begin{equation} \label{eq:signal-distributions-ordered}
x(i_j) \sim f_j, \quad j=1,\ldots,s,
\end{equation}
while the remaining $(p-s)$ observations have common density $f_0$, i.e., $x(i)\sim f_0$ for $i\not\in S$. 
We further assume that the observations $x$ are mutually independent.

We adopt now a Bayesian framework to measure statistical risks and assume that the underlying (unknown) support is {\em random}. Namely,
let the ordered support $P=(i_1,\ldots,i_s)$ have prior
\begin{equation} \label{eq:uniform-ordered}
\pi((i_1,\ldots, i_s)) = {(p-s)!}/{p!},
\end{equation}
for all distinct $1\le i_1, \ldots, i_s\le p$.
Consequently, the unordered support $S=\{i_1,\ldots,i_s\}$ is distributed uniformly over the collection of all sets of size $s$, with 
the unordered uniform distribution $\pi^{\text{u}}$. That is, for all $S\in\mathcal{S}:=\left\{S\subseteq\{1,\ldots,p\};|S|=s\right\}$, we have 
\begin{equation} \label{eq:uniform}
\pi^{\text{u}}
(\{i_1,\ldots, i_s\}) = \sum_{\sigma}\pi((i_{\sigma(1)},\ldots, i_{\sigma(s)})) = {(p-s)!s!}/{p!},
\end{equation}
where the sum is taken over all permuations of $\{1,2,\ldots,s\}$.

For any fixed configuration $P$, consider the loss function,
$$
\ell(\widehat{S},S) := \P[\widehat{S} \neq S] = \P_{P}[\widehat{S} \neq S],
$$
where the probability is taken over the randomness in the observations only. 
\stilian{(The $x(i)$'s are conditionally independent, given the ordered support $P$.)}{\fbox{added.  Agree?}}
The Bayes optimal procedures should minimize 
\begin{equation} \label{eq:Bayes-risk}
    \E_\pi \P[\widehat{S} \neq S],
\end{equation}
where the expectation is taken over the random configurations $P$, with a uniform distribution $\pi$ as specified in \eqref{eq:uniform-ordered}.

@@

The finite-sample optimality of the oracle thresholding procedure $\widehat{S}^*$ is intimately linked with the \emph{monotone likelihood ratio} (MLR) property.
\begin{definition}[Monotone Likelihood Ratio]
A family of positive densities on $\R$, $\{f_\delta, \delta \in U\}$, is said to have the MLR property if, for all $\delta_0, \delta_1\in U\subseteq\R$ such that $\delta_0 < \delta_1$, the likelihood ratio $\left(f_{\delta_1}(x)/f_{\delta_0}(x)\right)$ is an increasing function of $x$.
\end{definition}

The relationship between the above two notions is summarized next.

\begin{proposition} \label{prop:optimal-oracle-procedures}
Let the observations $x(i)$, $i=1,\ldots,p$ be as prescribed as in \eqref{eq:signal-distributions-ordered} through \eqref{eq:uniform-ordered}.
If each of $\{f_0, f_{1}\},\ldots,\{f_0,f_{s}\}$ form an MLR family, then the oracle data thresholding procedure $\widehat{S}^* = \{i\,|\, x(i)\ge x_{[s]}\}$ is finite-sample optimal in terms of Bayes risk $\E_\pi \P[\widehat{S} \neq S]$. That is,
\begin{equation} \label{eq:finite-sample-Bayes-optimal}
    \widehat{S}^* \in \argmin_{\widehat{S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$.
\end{proposition}

The proof of Proposition \ref{prop:optimal-oracle-procedures} is found in Section \ref{suppsec:proofs}.

We emphasize that the oracle thresholding procedures are in fact \emph{finite-sample optimal} in the above Bayesian context.
Further, our setup allows for different alternative distributions, and relaxes the assumptions of \citet{butucea2018variable} when studying distributional generalizations, where the alternatives are assumed to be identically distributed.

It remains to understand when the key MLR property holds. 
We elaborate on this question next.

\subsection{Bayes optimality under sub-exponential errors}
\label{subsec:optimal-procedure-sub-exponential}

Returning to the more concrete signal-plus-noise model \eqref{eq:model-additive}, it turns out that the error tail behavior is what determines the optimality of data thresholding procedures.
In this setting, log-concavity of the error densities is \emph{equivalent} to the MLR property (Lemma \ref{lemma:MLR-log-concavity}). This, in turn, yields the finite-sample optimality of data thresholding procedures (Proposition \ref{prop:log-concave}).

@@

\begin{proof}[Proof of Lemma \ref{lemma:MLR-log-concavity}]
Suppose MLR holds, we will show that $f_0(t) = \exp\{\phi(t)\}$ for some concave function $\phi$.
By the assumption of MLR, for any $x_1 < x_2$, setting $\delta_0 = 0$, and $\delta_1 = (x_2 - x_1)/2 > 0$, we have
\begin{equation*}
    \log{\frac{f_{\delta_1}(x_2)}{f_{\delta_0}(x_2)}}
    = \phi\left(\frac{(x_1+x_2)}{2}\right)- \phi(x_2) 
    \ge \phi(x_1)- \phi\left(\frac{(x_1+x_2)}{2}\right) 
    = \log{\frac{f_{\delta_1}(x_1)}{f_{\delta_0}(x_1)}}.
\end{equation*}
This implies that the log-density $\phi(t)$ is midpoint-concave, i.e., for all $x_1$ and $x_2$, we have,
\begin{equation}
    \phi\left(\frac{(x_1+x_2)}{2}\right) 
    \ge \frac{1}{2} \phi(x_1) + \frac{1}{2} \phi(x_2).
\end{equation}
For Lebesgue measurable functions, midpoint concavity is equivalent to concavity by the Sierpinki Theorem (see, e.g., Sec I.3 of \cite{donoghue2014distributions}). This proves the `only-if' part.

For the `if' part, when  $\phi(t) = \log{(f_0(t))}$ is log-concave, then for any $\delta_0 < \delta_1$, and any $x<y$, we have
\begin{equation} \label{eq:concavity-implies-MLR}
    \log{\frac{f_{\delta_1}(y)}{f_{\delta_0}(y)}} - \log{\frac{f_{\delta_1}(x)}{f_{\delta_0}(x)}}
    = \phi(y-\delta_1) - \phi(y-\delta_0) - \phi(x-\delta_1) + \phi(x-\delta_0) \ge 0,
\end{equation}
where the last inequality is a simple consequence of concavity (see Lemma \ref{lemma:four-point-concavity}). This proves the `if' part.
\end{proof}

@@

\begin{proof}[Proof of Theorem \ref{thm:necessary-strengthened}]
When errors are independent with log-concave density, the oracle thresholding procedure $\widehat{S}^*_p$, by Proposition \ref{prop:log-concave}, minimizes the Bayes risk \eqref{eq:Bayes-risk} among \emph{all} procedures. That is,
$$
\liminf_{p\to\infty} \inf_{\widehat{S}_p} \E_\pi\P[\widehat{S}_p\neq S_p]
\ge \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p].
$$
Since $\widehat{S}^*_p$ belongs to the class of all thresholding procedures, we have
\begin{align*}
    \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p] 
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \E_\pi\P[\widehat{S}_p\neq S_p] \\
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \inf_{S_p} \P[\widehat{S}_p\neq S_p] = 1,
\end{align*}
when $\overline{r}<g(\beta)$, where the last line follows from Theorem \ref{thm:necessary}.
\end{proof}

\begin{corollary}
\label{cor:minimax-over-dependence}
Let $D(F)$ be the collection of error arrays with common marginal $F$ as defined in \eqref{eq:common-marginal-distribution} where $F$ is an $\text{AGG}(\nu)$ distribution.
Let also $\widehat{S}^{\text{Bonf}}_p$ be Bonferroni's procedure as described in Theorem \ref{thm:sufficient}.
If $\underline{r}>g(\beta)$, then we have
\begin{equation} \label{eq:minimax-over-dependence-over}
    \limsup_{p\to\infty} \sup_{\substack{\mu\in\Theta_p^+(\beta,\underline{r})\\ {\cal E}\in D(F)}} \P(\widehat{S}^{\text{Bonf}}_p \neq S_p) = 0.
\end{equation}
Further, when $\underline{r}<g(\beta)$, and $F$ has a positive log-concave density $f$, we have
\begin{equation} \label{eq:minimax-over-dependence-under}
    \liminf_{p\to\infty} \inf_{\widehat{S}_p} \sup_{\substack{\mu\in\Theta_p^+(\beta,\underline{r})\\ {\cal E}\in D(F)}} \P(\widehat{S}_p \neq S_p) = 
    1,
\end{equation}
where the infimum on $\widehat{S}_p$ is taken over all procedures.
\end{corollary}

\begin{proof}[Proof of Corollary \ref{cor:minimax-over-dependence}]
Relation \eqref{eq:minimax-over-dependence-over} is a re-statement of Remark \ref{rmk:sufficient-strengthened}.

@@

The following result provides the general form of finite-sample Bayes optimal procedures.
It turns out that in general, \emph{likelihood ratio thresholding} is optimal.

\begin{proposition} \label{prop:likelihood-ratio-thresholding}
Let the observations $x(i)$, $i=1,\ldots,p$ have $s$ signals as prescribed in \eqref{eq:uniform-ordered} having common density $f_a$, and let the rest $(p-s)$ locations have common density $f_0$.
Define the likelihood ratios 
$$
L(i) := {f_a(x(i))}\big/{f_0(x(i))},
$$
and let $L_{[1]} \ge L_{[2]} \ge \ldots \ge L_{[p]}$ be the order statistics of the $L(i)$'s.
Then the procedure $\widehat{S}_{\text{opt}} = \{i\,|\,L(i) \ge L_{[s]}\}$ is finite-sample optimal in terms of Bayes risk. That is,
\begin{equation}
    \widehat{S}_{\text{opt}} \in \argmin_{\widehat{S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$, where the infimum on $\widehat{S}_p$ is taken over all procedures.
\end{proposition}

The proof of Proposition \ref{prop:likelihood-ratio-thresholding} is found  in Section \ref{suppsec:proofs}.

% \begin{remark}
% While the assumption that the alternatives have a common density $f_a$ is slightly more restrictive than in Proposition \ref{prop:optimal-oracle-procedures}, it is not excessively so if one considers a fully Bayesian alternative, where $f_a = p_1f_1 + p_2f_2 + \ldots + p_sf_s$ is a random mixture of alternatives $f_1,\ldots,f_s$ with probability weights $p_1,\ldots,p_s$.
% \end{remark}

@@

\begin{remark} \label{rmk:tail-assumptions}
Consider the model \eqref{eq:model-additive} with independent errors, Proposition \ref{prop:likelihood-ratio-thresholding}, and indeed, Example \ref{exmp:suboptimal-data-thresholding} demonstrate that thresholding procedures are in fact \emph{sub-optimal} for $\text{AGG}(\nu)$ models with $\nu<1$.
Therefore, the optimality of thresholding procedures (specifically, Bonferroni's procedure) only applies to $\text{AGG}(\nu)$ models with $\nu\ge1$. 
% This suggests that the condition $\nu\ge1$ in Corollary \ref{cor:minimax-over-dependence} is tight.

If we restrict the space of methods to only thresholding procedures, then results in Section \ref{subsec:point-wise-minimax} state that the phase transition phenomenon --- the 0-1 law in the sense of Corollary \ref{cor:point-wise-minimax} --- is universal in all error models with rapidly varying tails.
This includes $\text{AGG}(\nu)$ models {\it for all} $\nu>0$.
In contrast, models with heavy (regularly varying) tailed errors do not exhibit this phenomenon (see Theorem \ref{thm:heavy-tails}).
We summarize the properties of thresholding procedures in Table \ref{table:role-of-thresholding}. 
\end{remark}

>>>>

@@

\section{Bayes optimality in support recovery problems}
\label{subsec:Bayes-optimality}

In studying support recovery problems (e.g., \citet{arias2017distribution}), restrictions to the thresholding procedures are sometimes justified by arguing that such procedures are the ``reasonable'' choice for estimating the support set.
%  In Theorem \ref{thm:sufficient} we saw that a (FWER-controlling) thresholding procedure is indeed consistent for support recovery. 
% It is natural to ask if it will be sufficient to restrict our attention to only such procedures.
We show in this section that, perhaps surprisingly, for general error models, thresholding procedures are not always optimal, even when the observations are independent.

We shall identify the optimal procedure for support recovery problems under a Bayesian setting with general distributional assumptions (including but not limited to additive models \eqref{eq:model-additive}).
Specifically, we assume that there is an ordered set $P = (i_1,\ldots,i_s)$, $i_i\in\{1,\ldots,p\}$, and $s$ (not necessarily equal) densities $f_{1}, \ldots, f_{s}$, such that the observations indexed by $P$ have corresponding densities. That is,
\begin{equation} \label{eq:signal-distributions-ordered}
x(i_j) \sim f_j, \quad j=1,\ldots,s.
\end{equation}
Let also the rest $(p-s)$ observations have common density $f_0$, i.e., $x(i)\sim f_0$ for $i\not\in S$.
We further assume that the observations $x$ are mutually independent.

We adopt here a Bayesian framework to measure statistical risks. 
Let the ordered support $P=(i_1,\ldots,i_s)$ have prior
\begin{equation} \label{eq:uniform-ordered}
\pi((i_1,\ldots, i_s)) = {(p-s)!}/{p!},
\end{equation}
for all distinct $1\le i_1, \ldots, i_s\le p$.
Consequently, the unordered support $S=\{i_1,\ldots,i_s\}$ is distributed uniformly in the collection of all set of size $s$, with the unordered uniform distribution $\pi^{\text{u}}$. That is, for all for all $S\in\mathcal{S}:=\left\{S\subseteq\{1,\ldots,p\};|S|=s\right\}$, we have 
\begin{equation} \label{eq:uniform}
\pi^{\text{u}}
(\{i_1,\ldots, i_s\}) = \sum_{\sigma}\pi((i_{\sigma(1)},\ldots, i_{\sigma(s)})) = {(p-s)!s!}/{p!},
\end{equation}
where the sum is taken over all permuations of $\{1,2,\ldots,s\}$.

For any fixed configuration $P$, consider the loss function,
$$
\ell(\widehat{S},S) := \P[\widehat{S} \neq S] = \P_{P}[\widehat{S} \neq S],
$$
where the probability is taken over the randomness in the observations only.
The Bayes optimal procedures should minimize 
\begin{equation} \label{eq:Bayes-risk}
    \E_\pi \P[\widehat{S} \neq S],
\end{equation}
where the expectation is taken over the random configurations $P$, with a uniform distribution $\pi$ as specified in \eqref{eq:uniform-ordered}.

@@

\subsection{When is oracle thresholding Bayes optimal?} % under sub-exponential errors}
\label{subsec:optimal-procedure-sub-exponential}

An elegant answer of this question can be given in terms of the \emph{monotone likelihood ratio} (MLR) property.

\begin{definition}[Monotone Likelihood Ratio]
A family of positive densities on $\R$, $\{f_\delta, \delta \in U\}$, is said to have the MLR property if, for all $\delta_0, \delta_1\in U\subseteq\R$ such that $\delta_0 < \delta_1$, the likelihood ratio $\left(f_{\delta_1}(x)/f_{\delta_0}(x)\right)$ is an increasing function of $x$.
\end{definition}

The following result provides a general criterion for the finite-sample optimality of the oracle thresholding 
procedure $\widehat{S}^*$.

\begin{proposition} \label{prop:optimal-oracle-procedures}
Let the observations $x(i)$, $i=1,\ldots,p$ be as prescribed as in \eqref{eq:signal-distributions-ordered} through \eqref{eq:uniform-ordered}.
If each of $\{f_0, f_{1}\},\ldots,\{f_0,f_{s}\}$ form an MLR family, then the oracle data thresholding procedure $\widehat{S}^* = \{i\,|\, x(i)\ge x_{[s]}\}$ is finite-sample optimal in terms of Bayes risk $\E_\pi \P[\widehat{S} \neq S]$. That is,
\begin{equation} \label{eq:finite-sample-Bayes-optimal}
    \widehat{S}^* \in \argmin_{\widehat{S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$.
\end{proposition} 
\begin{proof}%[Proof of Proposition \ref{prop:optimal-oracle-procedures}]
The problem of support recovery can be equivalently stated as a classification problem, where the discrete parameter space is $\mathcal{S} = \{S\subseteq\{1,\ldots,p\}:|S|=s\}$, and the observation $x \in\R^p$ has likelihood $f(x|S)$ indexed by the support set $S$.

By the optimality of the Bayes classifier (see, e.g., \citep{domingos1997optimality}), a set estimator that maximizes the probability of support recovery is such that
$$
\widehat{S} \in \argmax_{S\in \mathcal{S}} f(x|S) \pi(S).
$$
Since we know from \eqref{eq:uniform} that $\pi(S)$ is uniform, the problem in our context reduces to showing that $f(x|\widehat{S}^*) = f(x|\widehat{S})$, where $f(x|S)$ is the conditional distribution of data given the unordered support $S$,
$$
f(x|S) 
= \sum_{P\in\sigma(S)} f(x|P) \pi^{\text{ord}}(P|S) 
= \frac{1}{s!} \left(\sum_{P\in\sigma{(S)}} \prod_{i=1}^s {f_{i}(x(P(i)))}\right) \prod_{k\not\in S}{f_0(x(k))},
$$
where $\sigma(S)$ is the set of all permutations of the indices in the support set $S$.

Suppose $\widehat{S} \neq \widehat{S}^*$, then there must be indices $j \in \widehat{S}$ and $j' \in \widehat{S}^c$ such that $x(j) \le x(j')$.
We exchange the labels of $x(j)$ and $x(j')$, and form a new estimate $\widehat{S}\,' = \big(\widehat{S}\setminus\{j\}\big)\cup\{j'\}$.
Comparing the likelihoods under $\widehat{S}$ and $\widehat{S}\,'$, we have
\begin{align}
    f(x|\widehat{S}) - f(x|\widehat{S}\,') 
    &= \frac{1}{s!} \sum_{P\in\sigma{(\widehat{S})}} \prod_{i=1}^s {f_{i}(x(P(i)))} f_0(x(j'))\prod_{k\not\in \widehat{S}\cup\{j'\}}{f_0(x(k))} - \nonumber \\
    &\quad\quad\quad - \frac{1}{s!} \sum_{P'\in\sigma{(\widehat{S}')}} \prod_{i=1}^s {f_{i}(x(P'(i)))} f_0(x(j)) \prod_{k\not\in \widehat{S}'\cup\{j\}}{f_0(x(k))} \nonumber \\
    &= \frac{1}{s!} \left(\sum_{i=1}^s a_i  \Big(f_i(x(j)) f_0(x(j')) - f_i(x(j')) f_0(x(j))\Big) \right) \prod_{k\not\in \widehat{S}\cup\{j'\}}{f_0(x(k))}, \label{eq:MLR-optimality-proof}
    % = \log{f_\delta(x(j))} + \log{f_0(x(j'))} - \log{f_\delta(x(j'))} - \log{f_0(x(j))} 
\end{align}
where the last equality follows by first summing over all permutations fixing $P(i) = j$ and $P'(i) = j'$, and setting $a_i = \sum_{P\in\sigma{(\widehat{S}\setminus\{j\})}} \prod_{i'\neq i} {f_{i'} (x(P(i')))}$. Notice that the $a_i$'s are non-negative.

Since $x(j) \le x(j')$, and that each of $\{f_0, f_{i}\}$ is an MLR family, we have
$$
\frac{f_i(x(j))}{f_0(x(j))} - \frac{f_i(x(j'))}{f_0(x(j'))} \le 0 \implies f_i(x(j)) f_0(x(j')) - f_i(x(j')) f_0(x(j)) \le 0.
$$
Using Relation \eqref{eq:MLR-optimality-proof}, we conclude that $f(x|\widehat{S}) \le f(x|\widehat{S}\,')$.
This implies that any estimator that is not $\widehat{S}^*$ may be improved, and the optimality follows.
\end{proof}

We emphasize that under the MLR conditions in Proposition \ref{prop:optimal-oracle-procedures},
 the oracle thresholding procedures are in fact \emph{finite-sample optimal} in the 
above Bayesian context. Further, our setup allows for different alternative distributions, and relaxes the assumptions 
of \citet{butucea2018variable} when studying distributional generalizations, where the alternatives are assumed to be identically distributed.

It remains to understand when the key MLR property holds.  We elaborate on this question next.
Returning to the more concrete signal-plus-noise model \eqref{eq:model-additive}, it turns out that the error tail behavior is what determines the optimality of data thresholding procedures.
In this setting, log-concavity of the error densities is \emph{equivalent} to the MLR property (Lemma \ref{lemma:MLR-log-concavity}). This, in turn, yields the finite-sample optimality of data thresholding procedures (Proposition \ref{prop:log-concave}).

@@

\begin{proof}%[Proof of Lemma \ref{lemma:MLR-log-concavity}]
Suppose MLR holds, we will show that $f_0(t) = \exp\{\phi(t)\}$ for some concave function $\phi$.
By the assumption of MLR, for any $x_1 < x_2$, setting $\delta_0 = 0$, and $\delta_1 = (x_2 - x_1)/2 > 0$, we have
\begin{equation*}
    \log{\frac{f_{\delta_1}(x_2)}{f_{\delta_0}(x_2)}}
    = \phi\left(\frac{(x_1+x_2)}{2}\right)- \phi(x_2) 
    \ge \phi(x_1)- \phi\left(\frac{(x_1+x_2)}{2}\right) 
    = \log{\frac{f_{\delta_1}(x_1)}{f_{\delta_0}(x_1)}}.
\end{equation*}
This implies that the log-density $\phi(t)$ is midpoint-concave, i.e., for all $x_1$ and $x_2$, we have,
\begin{equation}
    \phi\left(\frac{(x_1+x_2)}{2}\right) 
    \ge \frac{1}{2} \phi(x_1) + \frac{1}{2} \phi(x_2).
\end{equation}
For Lebesgue measurable functions, midpoint concavity is equivalent to concavity by the Sierpinki Theorem (see, e.g., Sec I.3 of \cite{donoghue2014distributions}). This proves the `only-if' part.

For the `if' part, when  $\phi(t) = \log{(f_0(t))}$ is log-concave, then for any $\delta_0 < \delta_1$, and any $x<y$, we have
\begin{equation} \label{eq:concavity-implies-MLR}
    \log{\frac{f_{\delta_1}(y)}{f_{\delta_0}(y)}} - \log{\frac{f_{\delta_1}(x)}{f_{\delta_0}(x)}}
    = \phi(y-\delta_1) - \phi(y-\delta_0) - \phi(x-\delta_1) + \phi(x-\delta_0) \ge 0,
\end{equation}
where the last inequality is a simple consequence of concavity (see Lemma \ref{lemma:four-point-concavity} below). This proves the `if' part.
\end{proof}
\begin{lemma} \label{lemma:four-point-concavity}
Let $\phi$ be any concave function on $\R$. For any $x<y\in\R$, and $\delta>0$ we have
$$
\phi(x) + \phi(y+\delta) \le \phi(y) + \phi(x+\delta).
$$
\end{lemma}

\begin{proof}%[Proof of Lemma \ref{lemma:four-point-concavity}]
Pick $\lambda = \delta/(y-x+\delta)$, by concavity of $f$ we have
\begin{equation} \label{eq:four-point-concavity-1}
    \lambda \phi(x) + (1-\lambda) \phi(y+\delta) 
    \le \phi(\lambda x + (1-\lambda)(y+\delta)) 
    = \phi(y),
\end{equation}
and
\begin{equation} \label{eq:four-point-concavity-2}
    (1-\lambda) \phi(x) + \lambda \phi(y+\delta)
    \le \phi((1-\lambda) x + \lambda(y+\delta)) 
    = \phi(x+\delta).
\end{equation}
Summing up \eqref{eq:four-point-concavity-1} and \eqref{eq:four-point-concavity-2} and we arrive at the conclusion as desired.
\end{proof}

@@

\begin{proof}%[Proof of Theorem \ref{thm:necessary-strengthened}]
When errors are independent with log-concave density, the oracle thresholding procedure $\widehat{S}^*_p$, by Proposition \ref{prop:log-concave}, minimizes the Bayes risk \eqref{eq:Bayes-risk} among \emph{all} procedures. That is,
$$
\liminf_{p\to\infty} \inf_{\widehat{S}_p} \E_\pi\P[\widehat{S}_p\neq S_p]
\ge \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p].
$$
Since $\widehat{S}^*_p$ belongs to the class of all thresholding procedures, we have
\begin{align*}
    \liminf_{p\to\infty} \E_\pi\P[\widehat{S}^*_p\neq S_p] 
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \E_\pi\P[\widehat{S}_p\neq S_p] \\
    &\ge \liminf_{p\to\infty} \inf_{\widehat{S}_p\in{\cal T}} \inf_{S_p} \P[\widehat{S}_p\neq S_p] = 1,
\end{align*}
when $\overline{r}<g(\beta)$, where the last line follows from Theorem \ref{thm:necessary}.
\end{proof}

@@

When the MLR condition in Proposition \ref{prop:optimal-oracle-procedures} is violated, the oracle thresholding 
procedures can in fact be sub-optimal (see Example \ref{exmp:suboptimal-data-thresholding}, below). 
The following result provides the general form of the finite-sample Bayes optimal procedures.
It turns out that in general, \emph{likelihood ratio thresholding} is optimal.

\begin{proposition} \label{prop:likelihood-ratio-thresholding}
Let the observations $x(i)$, $i=1,\ldots,p$ have $s$ signals as prescribed in \eqref{eq:uniform-ordered} having common density $f_a$, and let the rest $(p-s)$ locations have common density $f_0$.
Define the likelihood ratios 
$$
L(i) := {f_a(x(i))}\big/{f_0(x(i))},
$$
and let $L_{[1]} \ge L_{[2]} \ge \ldots \ge L_{[p]}$ be the order statistics of the $L(i)$'s.
Then the procedure $\widehat{S}_{\text{opt}} = \{i\,|\,L(i) \ge L_{[s]}\}$ is finite-sample optimal in terms of Bayes risk. That is,
\begin{equation}
    \widehat{S}_{\text{opt}} \in \argmin_{\widehat{S}} \E_\pi \P[\widehat{S} \neq S].
\end{equation}
for all $s$ and $p$, where the infimum on $\widehat{S}_p$ is taken over all procedures.
\end{proposition} 
\begin{proof}%[Proof of Proposition \ref{prop:likelihood-ratio-thresholding}]
The proof is entirely analogous to that of Proposition \ref{prop:optimal-oracle-procedures}.
Since we know from \eqref{eq:uniform} that $\pi(S)$ is uniform, the problem reduces to showing that $f(x|\widehat{S}_{\text{opt}}) = f(x|\widehat{S})$, where 
$$
\widehat{S} \in \argmax_{S \in \mathcal{S}} f(x|S) \pi(S).
$$
and
$f(x|S)$ is the conditional distribution of data given the unordered support $S$,
\begin{equation} \label{eq:likelihood-ratio-thresholding-proof}
    f(x|S) = \sum_P f(x|P) \pi^{\text{ord}}(P|S) = \prod_{j\in S} f_a(x(j)) \prod_{j\not\in S}{f_0(x(j))}.
\end{equation}
Suppose $\widehat{S} \neq \widehat{S}_{\text{opt}}$, then there must be indices $j \in \widehat{S}$ and $j' \in \widehat{S}^c$ such that $L(j) \le L(j')$.
If we exchange the labels of $L(j)$ and $L(j')$, that is, we form a new estimate $\widehat{S}\,' = \big(\widehat{S}\setminus\{j\}\big)\cup\{j'\}$,
comparing the log-likelihoods under $\widehat{S}$ and $\widehat{S}\,'$, we have
\begin{equation*}
    \log{f(x|\widehat{S})} - \log{f(x|\widehat{S}\,')} 
    = \log{f_a(x(j))} + \log{f_0(x(j'))} - \log{f_a(x(j'))} - \log{f_0(x(j))}.
\end{equation*}
By the definition of $L(j)$'s, and the order relations, we obtain
\begin{equation*}
    \log{f(x|\widehat{S})} - \log{f(x|\widehat{S}\,')} 
    = \log{L(j)} - \log{L(j')} \le 0
\end{equation*}
This implies that any estimator that is not $\widehat{S}_{\text{opt}}$ may be improved, and the optimality follows.
\end{proof}

@@

\subsection{Sub-optimality of data thresholding procedures}

We provide a slightly more general result on the sub-optimality of data thresholding procedures.

\begin{corollary} \label{cor:log-convex}
Consider the additive error model \eqref{eq:model-additive}.
Let the errors $\epsilon$ be independent with common distribution $F$.
Let each of the $s$ signals be located on $\{1,\ldots,p\}$ uniformly at random with equal magnitude $0<\delta<\infty$.
% , independently with probability $p_1,\ldots,p_s$.
% , {\color{red} so that we have $f_a = p_1f_1 + p_2f_2 + \ldots + p_sf_s$.}
If errors $\epsilon$ are iid with density $f$ log-convex on $[K, +\infty)$, then whenever $j\in\widehat{S}_{\text{opt}}$ for some $x(j) > K+\delta$, we must have $j'\in\widehat{S}_{\text{opt}}$ for all $j'$ such that $K+\delta \le x(j') < x(j)$.
\end{corollary}

Specifically, if there are $m$ observations exceeding $K+\delta_s$, with $m>s$, then the top $m-s$ observations will \emph{not} be included in the optimal estimator $\widehat{S}_{\text{opt}}$.
This shows that, in the case when the errors have super-exponential tails, the optimal procedures are in general \emph{not} data thresholding.

\begin{proof}[Proof of Corollary \ref{cor:log-convex}]
Since the alternative $f_{a}(t) = {f(t-\delta)}$ are log-convex on $[K+\delta, \infty)$, by Relation \eqref{eq:concavity-implies-MLR} in the proof of Lemma \ref{lemma:MLR-log-concavity} and appealing to log-convexity (rather than log-concavity), the likelihood ratio
$$
L(j) := \frac{f_a(x(j))}{f_0(x(j))} %= \sum_{i=1}^s \frac{p_if_{i}(x(j))}{f_0(x(j))},
$$
is decreasing in $x(j)$ on $[K+\delta, \infty)$.
The proof is complete by applying Proposition \ref{prop:likelihood-ratio-thresholding}.
\end{proof}

\begin{remark}[An open problem]
In the non-log-concave setting, where we know that thresholding procedures are suboptimal, likelihood thresholding procedures are promising, thanks to Proposition \ref{prop:likelihood-ratio-thresholding}.

In the case where the signals have difference sizes, however, likelihood thresholding procedures are undefined since 
in the proof of Proposition \ref{prop:likelihood-ratio-thresholding}, the signal densities need to be identical for Relation \eqref{eq:likelihood-ratio-thresholding-proof} to hold.  In such cases, the existence of an optimal procedure is an open problem.
\end{remark}

\section{Minimax optimality in exact support recovery}

We establish in this section minimax versions of our results from Chapter \ref{chap:exact-support-recovery}.
Specifically, if we restrict ourselves to \emph{the class of thresholding procedures} $\cal T$ (defined in \eqref{eq:thresholding-procedure}), then Bonferroni's procedure is minimax optimal, for \emph{any} fixed dependence structures in the URS class.
This is formalized in Corollary \ref{cor:point-wise-minimax} below.% in Section \ref{subsec:point-wise-minimax}.
We refer to this result as \emph{point-wise} minimax, to emphasize the fact that this optimality holds for every \emph{fixed} URS array.

Meanwhile, if we search over \emph{all procedures}, but expand the model space to include {\em all} dependence 
structures, then a different minimax optimality statement holds for Bonferroni's procedure.
This result, formally stated in Section \ref{subsec:minimax-over-dependence}, is a consequence of our characterization 
of the finite-sample Bayes optimality of thresholding procedures in Sections \ref{subsec:Bayes-optimality} and \ref{subsec:optimal-procedure-sub-exponential}.
% The result is of independent interest and it is the main contribution of this section.

\subsection{Point-wise minimax optimality for thresholding procedures}
\label{subsec:point-wise-minimax}

\begin{proof}%[Proof of Corollary \ref{cor:point-wise-minimax}]
The first conclusion \eqref{eq:point-wise-minimax-above} is a restatement of Theorem \ref{thm:sufficient}.
